---
title: "HDAT9800 2024 Individual Assessment 2"
subtitle: "Extended exploratory data analysis with visualisations"
author: "Maggie Burmeister"
date: "2024-07-01"
output:
  html_document:
    highlight: tango
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 14
    fig_height: 24
---

# Prepare

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install package

```{r install_load_package}
if (!requireNamespace("rvest", quietly = TRUE)) install.packages("rvest")
if (!requireNamespace("knitr", quietly = TRUE)) install.packages("knitr")
if (!requireNamespace("httr", quietly = TRUE)) install.packages("httr")
if (!requireNamespace("readr", quietly = TRUE)) install.packages("readr")
if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
if (!requireNamespace("Hmisc", quietly = TRUE)) install.packages("Hmisc")
if (!requireNamespace("details", quietly = TRUE)) install.packages("details")
if (!requireNamespace("paletteer", quietly = TRUE)) install.packages("paletteer", dependencies = TRUE)
if (!require(janitor)) install.packages("janitor", dependencies = TRUE)
if (!require(patchwork)) install.packages("patchwork", dependencies = TRUE)
library(httr)
library(readr)
library(dplyr)
library(janitor)
library(tidyverse)
library(knitr)
library(ggplot2)
library(patchwork)
library(Hmisc) #Library for describe data
library(details)
library(rvest)
library(rmarkdown)
library(rlang)
library(purrr)
library(gridExtra)
library(grid)
library(paletteer)
```
# Task 1

## Load and represent the dataset

```{r read_csv_data}
# Read data from the the csv file
zigong <- readr::read_csv("C:/Users/maggi/OneDrive/Documents/GitHub/individual-marked-assessment-2-Maggie1910/zigong/dat.csv")

#Represent data
head(zigong)
```

## Describe raw data
```{r summarise_raw_data}
summarise_raw_data <- zigong %>%
  describe()
details::details(
  summary = 'Click to show result!',
  summarise_raw_data
)
```

## Clean and transform dataframe

### Clean column names

#### Column name before clean
```{r column_name_original}
col_name_before <- names(zigong)
details::details(
  summary = 'Click to show results!',
  col_name_before
)
```
#### Column name after clean

-   Copy data frame to data_clean: this this case zigong to zigong_clean

```{r copy_dataframe}
# Copy data frame to data_clean
zigong_clean <- zigong %>%
  as_tibble()
```

-   Clean column  names

```{r column_name_clean}
zigong_clean <- zigong %>% 
  clean_names()

col_name_after <- names(zigong_clean)
details::details(
  summary = 'Click to show results!',
  col_name_after
)
```

### Converting categorical columns to factors

#### Review data frame before converting
```{r review_df_convert_data}
head(zigong_clean)
```

#### Convert and transform data

```{r columns_to_factorize}
# Convert relevant columns to factors with appropriate labels
# Function to factorize
convert_to_factor <- function(df, cols) {
  for (col in cols) {
    if (col %in% names(df)) {
      df[[col]] <- as.factor(df[[col]])
      print(paste("Column", col, "is converted to factor"))
    } else {
      print(paste("Columns", col, "does not exist in the data frame"))
    }
  }
  return(df)
}

# Column name list
categorial_columns <- c("destination_discharge", "admission_ward", "admission_way", "occupation", "discharge_department", "gender", "type_of_heart_failure", "nyha_cardiac_function_classification", "killip_grade", "type_ii_respiratory_failure", "consciousness", "respiratory_support", "outcome_during_hospitalization")

zigong_clean <- convert_to_factor(zigong_clean, categorial_columns)
```

### Review data frame after converting

```{r review df after converting}
str(zigong_clean)
head(zigong_clean)
```

### Renaming and Re-labeling the age_cat column

#### Review 'levels' columns age_cat before processing
```{r review_levels_age_cat_before}
levels(
  as.factor(
    zigong_clean[["age_cat"]]
    )
  )
```

#### Renaming and Re-labeling the age_cat column

```{r re-labeling_age-cat_column}
# Define proper labels for the age_cat column
age_labels <- c("21-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+")

# Relabel the age_cat column
zigong_clean <- zigong_clean %>%
  mutate(age_cat = factor(age_cat, labels = age_labels))
```

#### Review 'levels' age_cat column after processing

```{r review_levels_age_cat_after}
levels(
  as.factor(
    zigong_clean[["age_cat"]]
    )
  )
```

#### View data frame
```{r review_df_age_cat_after}
head(zigong_clean)
```

# Task 2 : Create Individual Plots

## Visualize and display plots

```{r viz}
# Age Distribution
p1 <- ggplot(zigong_clean, aes(x = age_cat)) +
  geom_bar(fill = "#B9DDF1") + 
  theme_minimal() +
  labs(title = "Age Group Distribution", x = "Age Group", y = "Count")

# Gender Distribution
p2 <- ggplot(zigong_clean, aes(x = gender)) + 
  geom_bar(fill = "#A5CFE9") +
  theme_minimal() +
  labs(title = "Gender Distribution", x = "Gender", y = "Count")

# Height Distribution
p3 <- ggplot(zigong_clean, aes(x = height)) +
  geom_bar(fill = "#9FCAE6") +
  theme_minimal() +
  labs(title = "Height Distribution", x = "Height (m)", y = "Count")

# Weight Distribution
p4 <- ggplot(zigong_clean, aes(x = weight)) +
  geom_histogram(binwidth = 5, fill = "#8DBBDC") +
  theme_minimal() +
  labs(title = "Weight Distribution", x = "Weight (kg)", y = "Count")

# BMI Distribution
p5 <- ggplot(zigong_clean, aes(x = bmi)) +
  geom_histogram(binwidth = 5, fill = "#78A9CE") +
  theme_minimal() +
  labs(title = "BMI Distribution", x = "BMI", y = "Count")

# Type of Heart Failure
p6 <- ggplot(zigong_clean, aes(x = type_of_heart_failure)) +
  geom_bar(fill = "#699AC2") +
  theme_minimal() +
  labs(title = "Type of Heart Failure Distribution", x = "Type of Heart Failure", y = "Count")

# Numeric variable 1: High Sensitivity Troponin
p7 <- ggplot(zigong_clean, aes(x = high_sensitivity_troponin)) +
  geom_histogram(binwidth = 0.1, fill = "#5F90BE") +
  scale_x_log10()+
  theme_minimal() +
  labs(title = "High Sensitivity Troponin Distribution", x = "HS Troponin", y = "Count")

# Numeric variable 2: Visit times
p8 <- ggplot(zigong_clean, aes(x = visit_times)) +
  geom_histogram(binwidth = 5, fill = "#4D7EAB") +
  theme_minimal() +
  labs(title = "Visit times Distribution", x = "Visit times", y = "Count")

# Numeric variable 3: Systolic blood pressure
p9 <- ggplot(zigong_clean, aes(x = systolic_blood_pressure)) +
  geom_histogram(binwidth = 5, fill = "#3F6E9A") +
  theme_minimal() +
  labs(title = "Systolic BP Distribution", x = "Systolic BP (mmHg)", y = "Count")

# Numeric variable 4: Diastolic blood pressure
p10 <- ggplot(zigong_clean, aes(x = diastolic_blood_pressure)) +
  geom_histogram(binwidth = 5, fill = "#2E5B88") +
  theme_minimal() +
  labs(title = "Diastolic BP Distribution", x = "Diastolic BP (mmHg)", y = "Count")
# Display plots
p1
p2
p3
p4
p5
p6
p7
p8
p9
p10
```

## Generate plots above into slides show

### Create list column_dict and function to display plots into slides

#### Create column dictionary

```{r col_viz_task2}
# List of col to plot
col_dict <- list(
  age_cat = "distribution",
  height = "distribution",
  weight = "distribution",
  high_sensitivity_troponin = "distribution",
  systolic_blood_pressure = "distribution",
  diastolic_blood_pressure = "distribution",
  type_of_heart_failure = "count",
  gender = "count",
  admission_ward = "count",
  occupation = "count"
)
```

#### Create function to display plots into slides

```{r col_viz_task2_function}

# Function to create plot
create_plot <- function(data, column, plot_type) {
  # Define ID column name
  column_id_unique = "inpatient_number"
  
  # The function creates a title with the first letter capitalized
  create_title <- function(col_name) {
    if (plot_type == "distribution") {
      paste("Distribution of", tools::toTitleCase(gsub("_", " ", col_name)))
    }
    else {
      paste("Count of", tools::toTitleCase(gsub("_", " ", col_name)))
    }
  }
  
  # Define base theme
  base_theme <- theme(
   plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
   axis.title = element_text(size = 8),
   axis.text = element_text(size = 6),
   legend.title = element_text(size = 8),
   legend.text = element_text(size = 6),
   legend.key.size = unit(0.5,"cm")
  )
  
  if (plot_type == "distribution") {
    # Distribution plot for continuous variables
    p <- ggplot(data, aes(x = !!sym(column), y = !!sym(column_id_unique))) +
      geom_point(fill = "#2C7E41", color ="#2C7E41") +
      ggtitle(create_title(column)) +
      coord_flip() +
      labs(x = tools::toTitleCase(gsub("_", " ", column)),
           y = tools::toTitleCase(gsub("_", " ",column_id_unique))) +
      theme_minimal() +
      base_theme
  } else if (plot_type == "count") {
    
    # Bar chart for categorical variables
    p <- ggplot(data, aes(x = !!sym(column), fill = !!sym(column))) +
      geom_bar() +
      geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) + 
      scale_fill_paletteer_d("lisa::AlbrenchDurer") +
      scale_fill_brewer(palette = "Blues") +
      ggtitle(create_title(column)) +
      coord_flip() +
      labs(x = tools::toTitleCase(gsub("_", " ", column)),
           y = "Count",
           fill = tools::toTitleCase(gsub("_", " ", column))) +
      theme(axis.text.y = element_text(angle = 0)) +
      theme_minimal() +
      base_theme
  }
  
  return(p)
}
```

### Display the slides
```{r display the slides}
# Use the function
plot_list <- map(names(col_dict), ~create_plot(zigong_clean, .x, col_dict[[.x]]))
# Display
plot_list
```

```{r viz_task2}

# Create a list of charts


# Add margin for subplot grid
margin_grid = theme(plot.margin = unit(c(0.5, 1, 1, 0.5), "cm"))
tg <- textGrob('Health Data Visualisation & Communication', gp = gpar(fontsize = 16, fontface = 'bold')) 
sg <- textGrob('Visualization Data of Hospitalized patients with heart failure: integrating electronic healthcare records and external outcome data', gp = gpar(fontsize = 10))
margin_title <- unit(0.5, "line")

# Grid each column plot
 grid.arrange(tg, sg, 
             arrangeGrob(grobs = lapply(plot_list, "+", margin_grid),
                         ncol = 2),
             heights = unit.c(grobHeight(tg) + 1.2*margin_title, 
                               grobHeight(sg) + margin_title, 
                               unit(1,"null")))
```

# Task 3

## Grouping slides using patchwork lib

For compare and contrast the graphs above, I will group as below:
Slide 1: Height, Weight, BMI (Height and Weight will have a positive correlation to BMI)
Slide 2: Age, Type of Heart Failure (I want to see if Age have any relation to Type of Heart Failure)
Slide 3: Weight, Type of Heart Failure (I want to see if Weight have any relation to Type of Heart Failure)
Slide 4: HS Troponin, Type of Heart Failure (I want to see if HS Troponin have any relation to Type of Heart Failure as HS Troponin is a biomarker that is used to detect cardiac injury, including myocardial infartion. Elevated levels of hs-troponin in the blood can indicate damage to the heart muscle. This biomarker is particularly sensitive and can detect even small amounts of cardiac injury )
Slide 5: Type of Heart Failure, Systolic Blood Pressure, Diastolic Blood Pressure (Blood pressure levels can influence the development, progression, and outcomes of heart failure)

```{r viz into small multiples}
#Combine plots into slide
slide1 <- p3 + p4 + p5 + plot_layout(ncol =3)
slide2 <- p1 + p6 + p8 + plot_layout(ncol =3)
slide3 <- p4 + p6 + plot_layout(ncol =2)
slide4 <- p6 + p7 + plot_layout(ncol =2)
slide5 <- p6 + p9 + p10 + plot_layout(ncol =3)

# Display slides
slide1
slide2
slide3
slide4
slide5
```

### Exporting Plots to PDF slides for Presentation
```{r save to pdf slides for presentation}
ggsave("slide1.pdf", slide1, width = 10, height = 5)
ggsave("slide2.pdf", slide2, width = 10, height = 5)
ggsave("slide3.pdf", slide3, width = 10, height = 5)
ggsave("slide4.pdf", slide4, width = 10, height = 5)
ggsave("slide5.pdf", slide5, width = 10, height = 5)
```

## Using ggplot2 facetting

```{r filter out non-defite values}
# Ensure the 'age' column is numeric
zigong_clean <- zigong_clean %>%
  mutate(age_cat = as.numeric(age_cat))
# Filter out non-finite values for the each plot
zigong_clean_age <- zigong_clean %>%
  filter(is.finite(age_cat))
zigong_clean_bmi <- zigong_clean %>%
  filter(is.finite(bmi))
zigong_clean_hs_troponin <- zigong_clean %>%
  filter(is.finite(high_sensitivity_troponin))
zigong_clean_systolic_bp <- zigong_clean %>%
  filter(is.finite(systolic_blood_pressure))
zigong_clean_diastolic_bp <- zigong_clean %>%
  filter(is.finite(diastolic_blood_pressure))
```

### Age Distribution by Type of Heart Failure

```{r age distribution by type_of_heart_failure}
# Viz the plot
slide2_1 <- ggplot(zigong_clean_age, aes(x = age_cat)) +
  geom_histogram(binwidth = 5, fill = "#81B1D6") +
  theme_minimal() +
  labs(title = "Age Distribution by Type of Heart Failure", x = "Age", y = "Count") +
  facet_wrap(~ type_of_heart_failure)

# Display graph
slide2_1
```

### BMI Distribytion by Type of Heart Failure 

```{r bmi distribution by type_of_heart_failure}
# Viz the plot
slide2_2 <- ggplot(zigong_clean_bmi, aes(x = bmi)) +
  geom_histogram(binwidth = 5, fill = "#5F90BE") +
  theme_minimal() +
  labs(title = "BMI Distribution by Type of Heart Failure", x = "BMI", y = "Count") +
  facet_wrap(~ type_of_heart_failure)

# Display graph
slide2_2
```

### HS Troponin Distribution by Type of Heart Failure

```{r hs_troponin distribution by type_of_heart_failure}
# Viz the plot
slide2_3 <- ggplot(zigong_clean_hs_troponin, aes(x = high_sensitivity_troponin)) +
  geom_histogram(binwidth = 5, fill = "#699AC2") +
  theme_minimal() +
  scale_x_log10() +
  labs(title = "HS Troponin Distribution by Type of Heart Failure", x = "HS Troponin", y = "Count") +
  facet_wrap(~ type_of_heart_failure)

# Display graph
slide2_3
```

### Systolic BP Distrinution by Type of Heart Failure

```{r systolic_bp distribution by type_of_heart_failure}
# Viz the plot
slide2_4 <- ggplot(zigong_clean_systolic_bp, aes(x = systolic_blood_pressure)) +
  geom_histogram(binwidth = 5, fill = "#78A9CE") +
  theme_minimal() +
  labs(title = "Systolic BP Distribution by Type of Heart Failure", x = "Systolic BP", y = "Count") +
  facet_wrap(~ type_of_heart_failure)

# Display graph
slide2_4
```

### Diastolic BP Distribution by Type of Heart Failure

```{r diastolic_bp distribution by type_of_heart_failure}
# Viz the plot
slide2_5 <- ggplot(zigong_clean_diastolic_bp, aes(x = diastolic_blood_pressure)) +
  geom_histogram(binwidth = 5, fill = "#4D7EAB") +
  theme_minimal() +
  labs(title = "Diastolic BP Distribution by Type of Heart Failure", x = "Diastolic BP", y = "Count") +
  facet_wrap(~ type_of_heart_failure)

# Display graph
slide2_5
```



