---
title: "HDAT9800 2024 Individual Assessment 1"
author: "Maggie Burmeister"
date: "2024-06-17"
output:
  html_document:
    highlight: tango
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
    fig_width: 14
    fig_height: 24
---

# Prepare

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Install Package

```{r install_load_package}
if (!requireNamespace("rvest", quietly = TRUE)) install.packages("rvest")
if (!requireNamespace("knitr", quietly = TRUE)) install.packages("knitr")
if (!requireNamespace("httr", quietly = TRUE)) install.packages("httr")
if (!requireNamespace("readr", quietly = TRUE)) install.packages("readr")
if (!requireNamespace("tidyverse", quietly = TRUE)) install.packages("tidyverse")
if (!requireNamespace("Hmisc", quietly = TRUE)) install.packages("Hmisc")
if (!requireNamespace("details", quietly = TRUE)) install.packages("details")
library(httr)
library(readr)
library(tidyverse)
library(knitr)
library(Hmisc) #Library for describe data
library(details)
library(rvest)
library(knitr)
library(rmarkdown)
library(rlang)
library(purrr)
library(gridExtra)
library(grid)
```

## Load and Export .Rdata

```{r load_from_Rdata}
data <- readRDS("data.rds")
data_clean <- readRDS("data_clean.rds")
```

```{r export_to_Rdata}
# Save a single object to a file
saveRDS(data, "data.rds")
saveRDS(data_clean, "data_clean.rds")
```

# Task

## Load data from link

### Load data

```{r read_data}
# URL of CSV file
data_url <- "https://figshare.com/ndownloader/files/35249488"

# Get data from URL
data_response <- GET(data_url)

# Checking errors
if (http_error(data_response)) {
  stop("Failed to download file")
}

# Read data from a response data
data <- read_csv(content(data_response, "text"))

# Represent data
head(data)
```

### Describe Raw Data

```{r summarise_raw_data}
summarise_raw_data <- data %>%
  describe()

details::details(
  summary = 'Click to show result!',
  summarise_raw_data
)
```

## Clean and transform dataframe

### Clean column name

#### Column name before clean

```{r column_name_original}
names(data)
```

#### Column name after clean

-   Copy dataframe to data_clean

```{r copy_dataframe}
# Copy dataframe to data_clean
data_clean <- data %>%
  as_tibble()
```

-   Clean columnn name

```{r column_name_clean}
# Use regex to clean " ", "-", "()" in column name
names(data_clean) <- gsub(" ", "", names(data))
names(data_clean) <- gsub("[()-]", "_", names(data_clean))
names(data_clean)
```

### Converting categorical columns to factors, with appropriate labels for each factor value

#### Create categorical dictionary

```{r categorical_dict}
# Dictionary for column categorical
category_dict <- list(
  Gender = c("1" = "Male", "2" = "Female"),
  Ethnic = c("1" = "Asian", "2" = "Afro", "3" = "Caucasian", "4" = "Other"),
  BaseDrugCombo = c("0" = "FTC + TDF", "1" = "3TC + АВС", "2" = "FTC + TAF", "3" = "DRV + FTC + TDF", "4" = "FTC + RTVB + TDF", "5" = "Other"),
  Comp.INI = c("0" = "DTG", "1" = "RAL", "2" = "EVG", "3" = "Not Applied"),
  Comp.NNRTI = c("0" = "NVP", "1" = "EFV", "2" = "RPV", "3" = "Not Applied"),
  ExtraPI = c("0" = "DRV", "1" = "RTVB", "2" = "LPV", "3" = "RTV", "4" = "ATV", "5" = "Not Applied"),
  Extrapk_En = c("0" = FALSE, "1" = TRUE),
  VL_M_ = c("0" = FALSE, "1" = TRUE),
  CD4_M_ = c("0" = FALSE, "1" = TRUE),
  Drug_M_ = c("0" = FALSE, "1" = TRUE)
)
```

#### Create function convert number to categorical label

```{r function_convert_categorical}
# Create function convert number to categorical label
map_to_labels <- function(column, dict) {
  return(factor(column, levels = names(dict), labels = dict))
}
```

#### Apply the function to categorical columns using a loop

```{r apply_func_categorical}
# Apply the function to categorical columns using a loop
for (col in names(category_dict)) {
  data_clean[[col]] <- map_to_labels(data_clean[[col]], category_dict[[col]])
  print(paste("Column '", col, "' is converted."))
}
```

#### Show dataframe with categorical columns is converted

-   Tibble data

```{r df_categorical_converted}
data_clean[1:100, ]
```

## Data Dictionary Table

```{r data_dictionary}
# Set the URL of the website that contains the table you want to get
url_dict_table <- "https://healthgym.ai/antiviral-hiv/"

# Read HTML from webpage
webpage <- read_html(url_dict_table)

# Get the table from the webpage
# Note: '.table' is an example of CSS selector.You can adjust it depend on
# the structure of a specific webpage.
table_data <- webpage %>% 
  html_node("table") %>% 
  html_table()

# Showing table in R Markdown
#table(table_data, caption = "Data Dictionary Table")
table_data
```

## Visualize the data of each column

```{r}
data_clean %>%
  select(PatientID, VL) %>%
  group_by(PatientID) %>%
  summarise(mean(VL, na.rm = TRUE))
```

```{r test_groupdata}
column_name = "VL"
data_clean %>%
  group_by(PatientID) %>%
  summarise(!!paste0(column_name, "_mean") := mean(!!sym(column_name), na.rm = TRUE))
```

```{r func_visualize_data}
# Function create plot for each column
create_plot <- function(data, column) {
  
  # Define ID column name
  ID_col_name = "PatientID"
  column_lst = c("PatientID", "...1", "Timepoints")
  
  # Define base theme
  base_theme <- theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
      axis.title = element_text(size = 8),
      axis.text = element_text(size = 6),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.5, "cm")
  )
  
  if (column %in% column_lst) {
    # With column ID, create a ggplot with text info
    plot_val <- ggplot() + 
      annotate("text", x = 0.5, y = 0.5, label = paste("There is no graph for the column ", column), size = 3) +
      ggtitle(paste("Visualization for ", column)) +
      theme_void() +
      theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 10))
    
  } else if (is.factor(data[[column]])) {
    # Bar chart for factor data
    plot_val <- ggplot(data, aes(x = !!sym(column), fill=!!sym(column))) +
      geom_bar() +
      scale_fill_brewer(palette="Blues") +
      ggtitle(paste("Visualization for", column)) +
      coord_flip() +
      labs(x = column, y = "Count") +
      theme(axis.text.y = element_text(angle = 0)) +
      theme_minimal() +
      base_theme
    
  } else {
    
    # Histogram chart for numeric data
    # Group the data by PatientID, averaging the data 
    # across observations across all patients.
    data_processed <- data %>%
      group_by(!!sym(ID_col_name)) %>%
      summarise(mean_value = mean(!!sym(column), na.rm = TRUE))
    
    # Visualization
    plot_val <- ggplot(data_processed, aes(x = mean_value)) +
      geom_histogram(aes(y=after_stat(density)), bins = 30, colour = "#FFFFFF", fill = "#9ECAE1") +
      geom_density(color = "blue") +
      ggtitle(paste("Visualization for", column)) +
      labs(x = column, y = "Frequency") +
      theme_minimal() +
      base_theme

  }
  return(plot_val)
}

# Create a list of charts
plot_list <- map(names(data_clean), ~create_plot(data_clean, .x))
```

```{r visualize_data, fig.width=14, fig.height=26}
# Arrange charts on a 2-column grid
# grid.arrange(grobs = plot_list, ncol = 2)

# Add margin for subplot grid
margin_grid = theme(plot.margin = unit(c(0.5, 1, 1, 0.5), "cm"))

tg <- textGrob('Health Data Visualisation & Communication', gp = gpar(fontsize = 16, fontface = 'bold'))
sg <- textGrob('Visualization Data of ART for HIV Antiretroviral Therapy for HIV (ART for HIV)', gp = gpar(fontsize = 10))
margin_title <- unit(0.5, "line")

# Grid each column plot


grid.arrange(tg, sg, 
             arrangeGrob(grobs = lapply(plot_list, "+", margin_grid),
                         ncol = 2),
             heights = unit.c(grobHeight(tg) + 1.2*margin_title, 
                              grobHeight(sg) + margin_title, 
                              unit(1,"null")))

```
